---
title: "Simulation setup merton model"
output: html_document
---

### Initialization

```{r, echo=FALSE}

# Clear variables
rm (list=ls())

# Settings for simulation grid
tFinal   <- 40                      # Final time --> naar 40 yr?
deltaT   <- 1                       # Time step
timeGrid <- seq(0, tFinal, deltaT)  # Time grid
nGrid    <- length(timeGrid)        # Time grid size
nSim     <- 10000                   # Number of simulated paths

# Settings for option formula and stochastic process
mu    <- 0.05 #is mu of risky asset
rfr   <- 0.01
sigma  <- 0.15 #is sigma of risky asset
s0     <- 1
gamma  <- 4.3

f_0 <- 1
h_t <- 0
m_max <- 1
mvt_bar <- (mu - rfr) / (sigma^2 * gamma)

jaar_salaris <- 30000 #input
franchise <- 17000
max_pensioengevend_salaris <- 137800
premie_afdracht <- 0.0175
premium_growth_rate <- 0.03
pensioengrondslag <- min(max_pensioengevend_salaris, jaar_salaris - franchise)
premie <- pensioengrondslag * premie_afdracht
```

### Simulate Brownian motion

```{r, echo=FALSE}
    
# Generate Brownian motion increments 
deltaW <- sqrt(deltaT) * matrix(rnorm(nSim/2*(nGrid-1)), ncol = nSim/2)

# Use antithetic sampling (variance reduction technique) 
deltaW <- cbind(deltaW,-deltaW)

# Compute Brownian motion paths
browMot <-  apply(deltaW, 2, cumsum)

# Add zeros as initial values
browMot <- rbind(integer(nSim),browMot)

# Simulate geometric Brownian motion (risk neutral)
geoBrowMot  <- s0*exp((mu - sigma^2/2) * timeGrid + sigma * browMot)

```

Above we have simulated geometric Brownian motion directly based on the analytically solution. We can also simulate geometric Brownian motion using a loop over time. This technique can later also be applied for more complex processes when no analytic solution is available.


#### Naar assignment toe
Eerst human capital bepalen: elk jaar premie h erbij. Impliciete aanname dat jaar/maand premieafdracht geen impact heeft op opbouw --> maandelijkse toekenning van rendement wordt genegeerd bij berekening van present values. 
```{r}
nSteps <- nGrid - 1
human_capital <- h_t*matrix(1, nGrid, 1)


for (iStep in 1:nSteps) {
  future_premiums <- matrix(0, nGrid - iStep, 1)
  
  for (iPremium in 1:(nGrid - iStep)) {
    future_premiums[iPremium, 1] <- premie * deltaT * exp(-(rfr - premium_growth_rate) * (iPremium - iStep))
  }
  
  if (iStep == 1) {
    print(future_premiums)
  }
  
  human_capital[iStep, 1] <- sum(future_premiums)
}

print(human_capital)
```



```{r}
financial_capital <- f_0*matrix(1, nGrid, nSim) #F
investment_fraction <- m_max*matrix(1, nGrid, nSim) #m_t
financial_fraction <- ((f_0 + human_capital[1,1])/f_0)*matrix(1, nGrid, nSim) #F + H / F

  #Doel F_(t+1) bepalen gegeven F_t -> daarna m_(t+1) bepalen
for(iStep in 1:nSteps){
  #F_t bepalen
  financial_capital[iStep+1,]  <- (1-investment_fraction[iStep, 1])* exp(rfr * deltaT)*financial_capital[iStep,] + investment_fraction[iStep, 1] * (geoBrowMot[iStep+1,] / geoBrowMot[iStep,]) * financial_capital[iStep,] + premie * deltaT * exp(rfr * deltaT)
  
  # Fractie (F_t + H_t) / F_t bepalen
  financial_fraction[iStep + 1,] <- (financial_capital[iStep+1,] + human_capital[iStep+1,]) / financial_capital[iStep+1,]
  
  # m_t bepalen
  investment_fraction[iStep + 1,] <- pmin(m_max, mvt_bar * financial_fraction[iStep + 1,])
}

```


```{r}
plot(rowMeans(investment_fraction))

cbind(Age=seq(68-tFinal, 68, by=1), Fraction_risky = rowMeans(investment_fraction))
```


```{r}
tFinal   <- 40                      # Final time --> naar 40 yr?
deltaT   <- 1                       # Time step
nSim     <- 10000                   # Number of simulated paths

# Settings for option formula and stochastic process
mu    <- 0.05 #is mu of risky asset
rfr   <- 0.01
sigma  <- 0.15 #is sigma of risky asset
s0     <- 1
gamma  <- 4.3

f_0 <- 1
h_t <- 0
m_max <- 1

jaar_salaris <- 30000 #input
franchise <- 17000
max_pensioengevend_salaris <- 137800
premie_afdracht <- 0.0175

get_lifecycle <- function(tFinal, deltaT, nSim, mu, rfr, sigma, s0, gamma, f_0, h_t, m_max, jaar_salaris, franchise, max_pensioengevend_salaris, premie_afdracht, premium_growth_rate){
  #inputs afmaken
  timeGrid <- seq(0, tFinal, deltaT)
  nGrid <- length(timeGrid)
  mvt_bar <- (mu - rfr) / (sigma^2 * gamma)
  pensioengrondslag <- min(max_pensioengevend_salaris, jaar_salaris - franchise)
  premie <- pensioengrondslag * premie_afdracht
  
  ########################################
  # Generate Brownian motion increments ##
  ########################################
  deltaW <- sqrt(deltaT) * matrix(rnorm(nSim/2*(nGrid-1)), ncol = nSim/2)

  # Use antithetic sampling (variance reduction technique) 
  deltaW <- cbind(deltaW,-deltaW)

  # Compute Brownian motion paths
  browMot <-  apply(deltaW, 2, cumsum)

  # Add zeros as initial values
  browMot <- rbind(integer(nSim),browMot)

  # Simulate geometric Brownian motion (risk neutral)
  geoBrowMot  <- s0*exp((mu - sigma^2/2) * timeGrid + sigma * browMot)
  
  ################################
  ### Generate human capital ####
  ##############################
  nSteps <- nGrid - 1
  human_capital <- h_t*matrix(1, nGrid, 1)

  
for (iStep in 1:nSteps) {
  future_premiums <- matrix(0, nGrid - iStep, 1)
  
  for (iPremium in 1:(nGrid - iStep)) {
    future_premiums[iPremium, 1] <- premie * deltaT * exp(-(rfr - premium_growth_rate)* (iPremium - iStep))
  }
  
  human_capital[iStep, 1] <- sum(future_premiums)
}
  
  #####################
  #### Financial ######
  #####################
  financial_capital <- f_0*matrix(1, nGrid, nSim) #F
investment_fraction <- m_max*matrix(1, nGrid, nSim) #m_t
financial_fraction <- ((f_0 + human_capital[1,1])/f_0)*matrix(1, nGrid, nSim) #F + H / F

  #Doel F_(t+1) bepalen gegeven F_t -> daarna m_(t+1) bepalen
for(iStep in 1:nSteps){
  #F_t bepalen
  financial_capital[iStep+1,]  <- (1-investment_fraction[iStep, 1])* exp(rfr * deltaT)*financial_capital[iStep,] + investment_fraction[iStep, 1] * (geoBrowMot[iStep+1,] / geoBrowMot[iStep,]) * financial_capital[iStep,] + premie * deltaT * exp(rfr * deltaT)
  
  # Fractie (F_t + H_t) / F_t bepalen
  financial_fraction[iStep + 1,] <- (financial_capital[iStep+1,] + human_capital[iStep+1,]) / financial_capital[iStep+1,]
  
  # m_t bepalen
  investment_fraction[iStep + 1,] <- pmin(m_max, mvt_bar * financial_fraction[iStep + 1,])
}

lifeCycle <- rowMeans(investment_fraction)

#### resultaten maken #####
resulten_list <- list(human_capital = human_capital, 
                      financial_capital = financial_capital, 
                      financial_fraction = financial_fraction, 
                      investment_fraction = investment_fraction, 
                      lifecycle = lifeCycle)
  
return(resulten_list)
}
```



```{r}
gamma_vector <-  c(4.6,	3.1,	2.1)
horizon <- 40
res_lifecycles <- data.frame(age=c(), m_t=c(), lifecycle=c())

for (i in 1:length(gamma_vector)){
 temp <- get_lifecycle(tFinal = horizon, 
              deltaT = 1, 
              nSim = 10000, 
              mu = 0.053, 
              rfr = 0.025, 
              sigma = 0.173, 
              s0 = 1, 
              gamma = gamma_vector[i], 
              f_0 = 1, 
              h_t = 0,
              m_max = 1,
              jaar_salaris = 30000, 
              franchise = 17000,
              max_pensioengevend_salaris = 137800, 
              premie_afdracht = 0.0175,
              premium_growth_rate = 0.03)
  
 res_lifecycles <- rbind(res_lifecycles, data.frame(age=(68-horizon):68, m_t=temp$lifecycle, lifecycle=toString(gamma_vector[i])))
}

head(res_lifecycles)

library(ggplot2)
ggplot(res_lifecycles) + 
  geom_line(aes(x=age,y=m_t, linetype = lifecycle), lwd=1)+ #can ook linetype vervangen door colour
  labs(title = "Share of financial wealth invested in risky assets", y= "Share", x = "Age", linetype="Risk aversion (γ)") + 
  ylim(c(0,1))+
  theme_minimal()

```




```{r}
get_lifecycle_switch <- function(tFinal, deltaT, nSim, mu, rfr, sigma, s0, gamma, f_0, h_t, m_max, jaar_salaris, franchise, max_pensioengevend_salaris, premie_afdracht, premium_growth_rate, gamma_end, switch_x_years_before_end){
  #inputs afmaken
  timeGrid <- seq(0, tFinal, deltaT)
  nGrid <- length(timeGrid)
  pensioengrondslag <- min(max_pensioengevend_salaris, jaar_salaris - franchise)
  premie <- pensioengrondslag * premie_afdracht
  
  ########################################
  # Generate Brownian motion increments ##
  ########################################
  deltaW <- sqrt(deltaT) * matrix(rnorm(nSim/2*(nGrid-1)), ncol = nSim/2)

  # Use antithetic sampling (variance reduction technique) 
  deltaW <- cbind(deltaW,-deltaW)

  # Compute Brownian motion paths
  browMot <-  apply(deltaW, 2, cumsum)

  # Add zeros as initial values
  browMot <- rbind(integer(nSim),browMot)

  # Simulate geometric Brownian motion (risk neutral)
  geoBrowMot  <- s0*exp((mu - sigma^2/2) * timeGrid + sigma * browMot)
  
  ################################
  ### Generate human capital ####
  ##############################
  nSteps <- nGrid - 1
  human_capital <- h_t*matrix(1, nGrid, 1)

  
for (iStep in 1:nSteps) {
  future_premiums <- matrix(0, nGrid - iStep, 1)
  
  for (iPremium in 1:(nGrid - iStep)) {
    future_premiums[iPremium, 1] <- premie * deltaT * exp(-(rfr - premium_growth_rate)* (iPremium - iStep))
  }
  
  human_capital[iStep, 1] <- sum(future_premiums)
}
  
  #####################
  #### Financial ######
  #####################
  financial_capital <- f_0*matrix(1, nGrid, nSim) #F
investment_fraction <- m_max*matrix(1, nGrid, nSim) #m_t
financial_fraction <- ((f_0 + human_capital[1,1])/f_0)*matrix(1, nGrid, nSim) #F + H / F

  #Doel F_(t+1) bepalen gegeven F_t -> daarna m_(t+1) bepalen
for(iStep in 1:nSteps){
  #F_t bepalen
  financial_capital[iStep+1,]  <- (1-investment_fraction[iStep, 1])* exp(rfr * deltaT)*financial_capital[iStep,] + investment_fraction[iStep, 1] * (geoBrowMot[iStep+1,] / geoBrowMot[iStep,]) * financial_capital[iStep,] + premie * deltaT * exp(rfr * deltaT)
  
  # Fractie (F_t + H_t) / F_t bepalen
  financial_fraction[iStep + 1,] <- (financial_capital[iStep+1,] + human_capital[iStep+1,]) / financial_capital[iStep+1,]
  
  #mvt_bar updaten afhankelijk van periode
  gamma_used <- ifelse(iStep <= tFinal - switch_x_years_before_end , gamma, gamma_end)
  
  print(gamma_used)
  mvt_bar <- (mu - rfr) / (sigma^2 * gamma_used)
  
  # m_t bepalen
  investment_fraction[iStep + 1,] <- pmin(m_max, mvt_bar * financial_fraction[iStep + 1,])
}

lifeCycle <- rowMeans(investment_fraction)

#### resultaten maken #####
resulten_list <- list(human_capital = human_capital, 
                      financial_capital = financial_capital, 
                      financial_fraction = financial_fraction, 
                      investment_fraction = investment_fraction, 
                      lifecycle = lifeCycle)
  
return(resulten_list)
}
```

```{r}
gamma_vector <- c(2, 2.5, 3, 4, 8)
horizon <- 40
res_lifecycles_switch <- data.frame(age=c(), m_t=c(), lifecycle=c())

for (i in 1:length(gamma_vector)){
  print(paste("new gamma of value ", gamma_vector[i], sep = ""))
 temp <- get_lifecycle_switch(tFinal = horizon, 
              deltaT = 1, 
              nSim = 10000, 
              mu = 0.05, 
              rfr = 0.01, 
              sigma = 0.15, 
              s0 = 1, 
              gamma = gamma_vector[i], 
              f_0 = 1, 
              h_t = 0,
              m_max = 1,
              jaar_salaris = 30000, 
              franchise = 17000,
              max_pensioengevend_salaris = 137800, 
              premie_afdracht = 0.0175,
              premium_growth_rate = 0.03,
              gamma_end = gamma_vector[3],
              switch_x_years_before_end = 10)
  
 res_lifecycles_switch <- rbind(res_lifecycles_switch, data.frame(age=(68-horizon):68, m_t=temp$lifecycle, lifecycle=paste("gamma", gamma_vector[i], sep = "_")))
}

head(res_lifecycles_switch)
ggplot(res_lifecycles_switch) + 
  geom_line(aes(x=age,y=m_t, linetype = lifecycle), lwd=1)+ #can ook linetype vervangen door colour
  labs(title = "Share of financial wealth invested in risky assets", y= "Share", x = "Age", linetype="Risk aversion") + 
  theme_minimal()
```




### Ingroeien naar default

```{r}
get_lifecycle_ingroeien <- function(tFinal, deltaT, nSim, mu, rfr, sigma, s0, gamma, f_0, h_t, m_max, jaar_salaris, franchise, max_pensioengevend_salaris, premie_afdracht, premium_growth_rate, gamma_end, switch_x_years_before_end){
  #inputs afmaken
  timeGrid <- seq(0, tFinal, deltaT)
  nGrid <- length(timeGrid)
  pensioengrondslag <- min(max_pensioengevend_salaris, jaar_salaris - franchise)
  premie <- pensioengrondslag * premie_afdracht
  
  ########################################
  # Generate Brownian motion increments ##
  ########################################
  deltaW <- sqrt(deltaT) * matrix(rnorm(nSim/2*(nGrid-1)), ncol = nSim/2)

  # Use antithetic sampling (variance reduction technique) 
  deltaW <- cbind(deltaW,-deltaW)

  # Compute Brownian motion paths
  browMot <-  apply(deltaW, 2, cumsum)

  # Add zeros as initial values
  browMot <- rbind(integer(nSim),browMot)

  # Simulate geometric Brownian motion (risk neutral)
  geoBrowMot  <- s0*exp((mu - sigma^2/2) * timeGrid + sigma * browMot)
  
  ################################
  ### Generate human capital ####
  ##############################
  nSteps <- nGrid - 1
  human_capital <- h_t*matrix(1, nGrid, 1)

  
for (iStep in 1:nSteps) {
  future_premiums <- matrix(0, nGrid - iStep, 1)
  
  for (iPremium in 1:(nGrid - iStep)) {
    future_premiums[iPremium, 1] <- premie * deltaT * exp(-(rfr - premium_growth_rate)* (iPremium - iStep))
  }
  
  human_capital[iStep, 1] <- sum(future_premiums)
}
  
  #####################
  #### Financial ######
  #####################
  financial_capital <- f_0*matrix(1, nGrid, nSim) #F
investment_fraction <- m_max*matrix(1, nGrid, nSim) #m_t
financial_fraction <- ((f_0 + human_capital[1,1])/f_0)*matrix(1, nGrid, nSim) #F + H / F

  #Doel F_(t+1) bepalen gegeven F_t -> daarna m_(t+1) bepalen
for(iStep in 1:nSteps){
  #F_t bepalen
  financial_capital[iStep+1,]  <- (1-investment_fraction[iStep, 1])* exp(rfr * deltaT)*financial_capital[iStep,] + investment_fraction[iStep, 1] * (geoBrowMot[iStep+1,] / geoBrowMot[iStep,]) * financial_capital[iStep,] + premie * deltaT * exp(rfr * deltaT)
  
  # Fractie (F_t + H_t) / F_t bepalen
  financial_fraction[iStep + 1,] <- (financial_capital[iStep+1,] + human_capital[iStep+1,]) / financial_capital[iStep+1,]
  
  #mvt_bar updaten afhankelijk van periode
  ratio <- max(0,(iStep - tFinal + switch_x_years_before_end))/switch_x_years_before_end
  gamma_used <- gamma*(1-ratio) + gamma_end*ratio
  # print(gamma_used)
  
  mvt_bar <- (mu - rfr) / (sigma^2 * gamma_used)
  
  # m_t bepalen
  investment_fraction[iStep + 1,] <- pmin(m_max, mvt_bar * financial_fraction[iStep + 1,])
}

lifeCycle <- rowMeans(investment_fraction)

#### resultaten maken #####
resulten_list <- list(human_capital = human_capital, 
                      financial_capital = financial_capital, 
                      financial_fraction = financial_fraction, 
                      investment_fraction = investment_fraction, 
                      lifecycle = lifeCycle)
  
return(resulten_list)
}
```

```{r}
gamma_vector <- c(4.6,	3.1,	2.1)
horizon <- 40
res_lifecycles_ingroeien <- data.frame(age=c(), m_t=c(), lifecycle=c())

for (i in 1:length(gamma_vector)){
  print(paste("new gamma of value ", gamma_vector[i], sep = ""))
 temp <- get_lifecycle_ingroeien(tFinal = horizon, 
              deltaT = 1, 
              nSim = 10000, 
              mu = 0.053, 
              rfr = 0.025, 
              sigma = 0.173, 
              s0 = 1, 
              gamma = gamma_vector[i], 
              f_0 = 1, 
              h_t = 0,
              m_max = 1,
              jaar_salaris = 30000, 
              franchise = 17000,
              max_pensioengevend_salaris = 137800, 
              premie_afdracht = 0.0175,
              premium_growth_rate = 0.03,
              gamma_end = 4.1,
              switch_x_years_before_end = 10)
  
 res_lifecycles_ingroeien <- rbind(res_lifecycles_ingroeien, data.frame(age=(68-horizon):68, m_t=temp$lifecycle, lifecycle=toString(gamma_vector[i])))
}

head(res_lifecycles_ingroeien)
ggplot(res_lifecycles_ingroeien) + 
  geom_line(aes(x=age,y=m_t, linetype = lifecycle), lwd=1)+ #can ook linetype vervangen door colour
  labs(title = "Share of financial wealth invested in risky assets", y= "Share", x = "Age", linetype="Risk aversion (γ)") + 
  ylim(c(0,1))+
  theme_minimal()
```



